Methods with preconditioning with weight decay regularization

### 2023-04-22 paper/main.tex
edit paper

\documentclass{article}
\usepackage{arxiv}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\author{ Kreinin M. \\
	Department of Intelligent Systems\\
	MIPT\\
	Moscow, Russia \\
	\texttt{kreinin.mv@phystech.edu} \\
	\And
	Beznosikov A. \\
	Department of Data Science\\
	MIPT\\
	Dolgoprudny, Russia \\
	\texttt{beznosikov.an@phystech.edu} \\
}
\date{}

\renewcommand{\shorttitle}{\textit{arXiv} Template}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={A template for the arxiv style},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={Kreinin M.},
pdfkeywords={First keyword, Second keyword, More},
}

@@

\begin{abstract}
	This paper examines the convergence rate of adaptive gradient methods when a regularization function is added to the target function. This is an area of research, since many machine learning problems use the heuristic, heuristic regularization, and we investigate the theoretical and practical convergence of adaptive gradient methods.
\end{abstract}

\keywords{Adam \and OASIS \and Regularization \and ADAHessian \and Weight Decay}

\section{Introduction}
In machine learning we consider unconstrained optimization problem
\begin{equation}
	\min_{w \in \mathbb{R}^d} f(w)
\end{equation}
Problems of the form (1) cover a plethora of applications, including empirical risk minimization,
deep learning \citep{lecun2015deep}, and supervised learning \citep{cunningham2008supervised} tasks such as regularized least squares \citep{rifkin2007notes} or logistic regression \citep{shalev2014understanding}. This minimization problem can be difficult to solve, particularly when the number of training
samples n, or problem dimension d, is large, or if the problem is nonconvex.

Second order preconditioners methods are methods when instead of calculate true hessian of our function for newton method \citep{dennis1977quasi}, some heuristics is used to calculate pseudo-hessian. There are second order preconditioners methods for solving this problem, such as Adam \citep{kingma2014adam} , OASIS \citep{goldberg2011oasis}, AdaHessian \citep{yao2021adahessian}.

Throughout this work we assume that each f $:\mathbb{R}^d \rightarrow \mathbb{R}$ is twice differentiable and also L-smooth. This is formalized in the following assumption.

\begin{assumption}{1}{Convex}
	The function f is convex, i.e. $\forall w, w' \in \mathbb{R}^d$
	\begin{equation}
		f(w) \geq f(w') + \langle \nabla F(w'), w-w' \rangle
	\end{equation}
\end{assumption}

\begin{assumption}{2}{L-smoothness}
	The gradients of F are L-Lipschitz continuous $\forall w \in \mathbb{R}^d$, i.e. there exists a constant $L > 0$ such that $\forall w, w' \in \mathbb{R}^d$,
	\begin{equation*}
		f(w) \leq f(w') + \langle \nabla f(w'), w-w' \rangle + \frac{L}{2} ||w - w'||^2
	\end{equation*}
\end{assumption}

\begin{assumption}{3}{Twice differentiable}
	The function f is twice continuously differentiable.
\end{assumption}

\begin{assumption}{4}{$\mu$ - strongly convex}
	The function f is $\mu$-strongly convex, i.e., there exists a constant $\mu > 0$ such that $\forall w, w' \in \mathbb{R}^d$
	\begin{equation*}
		f(w) \geq f(w') + \langle f(w'), w-w' \rangle +\frac{\mu}{2} ||w - w'||^2
	\end{equation*}

\end{assumption}

\section{Problem statement}
We want to investigate the convergence speed of the AdamW method and the newly proposed MyAdamW method in machine learning problems, and we also plan to prove the convergence of these methods and investigate the obtained solution.

We consider the unconstrained optimization problem
\begin{equation*}
	\min_{w \in \mathbb{R}^d} f(w)
\end{equation*}

But we can add regulirazation $r(w)$ -- regularization, and solve the unconstrained optimization problem.
\begin{equation*}
	\min_{w \in \mathbb{R}^d} F(w) = f(w) + r(w)
\end{equation*}

In the convergence algorithm that we study, we will investigate algorithms of the following two kinds. In the first one, the regularization function is taken out separately in the recalculation of model weights, and in the second one, the function is dominated by the inverse "hessian".

\begin{algorithm}[H]
	\caption{Based algorithm}\label{alg:Based}
	$w_0, D_0, \eta_0$
	\begin{algorithmic}
		\For{$k = 1, 2, ...$}
		\State $D_k$ - update by using information of $f(w)$
		\State $\eta_k$ - update
		\State Set $w_{k+1} = w_k - \eta_k D_k^{-1} \nabla f(w_k) - \eta \nabla r(w)$ 
		\State or like that
		\State Set $w_{k+1} = w_k - \eta_k D_k^{-1} (\nabla f(w_k) + \nabla r(w))$
		\EndFor
	\end{algorithmic}
\end{algorithm}

\section{Experiment}
% Мы проведем эксперимент, в котором сравним Adam с добавленной L2 регуляризацией с AdamW, при различных learning rate и decoupled weight, также мы сравним его с измененным алгоритмом AdamW, где регуляризатор занесен под гессиан, также мы сравним его при условии, что теперь минимальное собственное число гессиана, это не epsilon, а гиперпараметр decoupled weight.Проведем те же эксперименты со SGD и L2 регуляризацией и SGD с decoupled weitght при тех же параметрах. Все эксперименты будут проводиться на батчах размера 128, на сетке ResNet18, в 200 эпох. Всё это будет реализовано с помощью библиотеки PyTorch на dataset CIFAR10.

We will conduct an experiment in which we compare Adam with added L2 regularization with AdamW, at different learning rate and decoupled weight, also we will compare it with the modified AdamW algorithm where the regularizer is put under the hessian, also we will compare it under the condition that now the minimal eigenvalue of hessian is not epsilon, but decoupled weight hyperparameter.We will conduct the same experiments with SGD and L2 regularization and SGD with decoupled weitght at the same parameters. All experiments will be performed on batches of size 128, on a ResNet18 grid of 200 epochs. All this will be implemented using the PyTorch library on dataset CIFAR10.

\section{Theory}
To proof this algorigthm we use simplified form of AdamW algorithm
\begin{algorithm}[H]
\caption{AdamW}\label{alg:AdamW}

    \begin{algorithmic}
    \Require{$r, \varepsilon, f$}    
    \While {$w$ not converged}
        \State $t = t+1$
        \State $D_t = \text{diag}(|\nabla f(w_t)|_i)$
        \State $w_t = w_{t-1} - \eta \cdot \nabla f(w_t) D_t^{-1}  -\lambda \nabla r(w_t)$ 
    \EndWhile
\end{algorithmic}
\end{algorithm}

\begin{theorem}

Suppose the Assumption 1, 2, 5 and let $\varepsilon > 0$ and let the step-size satisfy
\begin{equation*}
    \eta < \frac{2 \alpha}{L + l \cdot \alpha} 
\end{equation*}
Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem (link here to problem 1) can be bounded by
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}

@@

Let's write first assumption for step t and $t+1$:

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \nabla f(w_t), w_{t+1} - w_t \rangle + \frac{L}{2}||w_{t+1} - w_t ||^2
\end{equation*}
Okay, by definition for our algorithm we have:

\begin{equation*}
w_{t+1} - w_t = -\eta D_t^{-1} \nabla f(w_t) - \eta \nabla r(w_t)    
\end{equation*}

and

\begin{equation*}
\nabla f(w_t) = \frac{1}{\eta} D^t(w_t - w_{t+1}) - D^t \nabla r(w_t)
\end{equation*}

Okay, now let's replace $\nabla f(w_t)$ and $I \leq \frac{D_t}{\alpha}$
\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \frac{1}{\eta}D_t(w_t - w_{t+1}) - D_t\nabla r(w_t), w_{t+1} - w_t \rangle + \frac{L}{2 \alpha} ||w_{t+1} - w_t||_{D_t}^2
\end{equation*}

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left(\frac{L}{2 \alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 - \langle D_t \nabla r(w_t), w_{t+1} - w_t \rangle
\end{equation*}

Lets define new variable $\tilde{r} : \nabla \tilde{r} = D_t \nabla r(w_t)$. Then rewrite step using the variable and 5-th assumption.
\begin{equation*}
    \tilde{r}(w_{t+1}) \leq \tilde{r}(w_t) + \langle \tilde{r}(w_t), w_{t+1} - w_t \rangle + \frac{l}{2} (w_{t+1} - w_t)^T D_t (w_{t+1} - w_t)
\end{equation*}

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left( \frac{L}{2\alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 + \tilde{r}(w_t) - \tilde{r}(w_{t+1}) + \frac{l}{2}||w_{t+1}-w_t||_{D_t}^2
\end{equation*}

$\tilde{F}(w) = f(w) + \tilde{r}(w)$, $F(w) = f(w) + r(w)$, ($\tilde{L}=L + l \alpha$), we get:

>>>>

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[english]{babel}	% локализация и переносы
%\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}		% Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}

\graphicspath{{pictures/}, {images/}, {}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% Цвета 
\usepackage[dvipsnames]{xcolor}
\usepackage{color}
%\usepackage{colortbl}

@@

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{Kreinin M. \\
	Department of Data Science\\
	Moscow Institute of Physics and Technology\\
	Moscow, Russia \\
	\texttt{kreinin.mv@phystech.edu} \\
	%% examples of more authors
        \And
        Beznosikov A. \\
	Department of Data Science\\
	Moscow Institute of Physics and Technology\\
	Moscow, Russia \\
	\texttt{beznosikov.an@phystech.edu} \\
	%% \AND
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{\textit{arXiv} Template}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf

@@

\begin{abstract}
This paper investigates the convergence behavior of optimization methods that utilize weight decay regularization, specifically focusing on popular variants such as AdamW.
We explore different alternatives to these method, with the goal of investigating their convergence speed and accuracy.
Also we conduct experiments on benchmark datasets and models in order to compare them on practice.
Overall, our study provides insights into the design of regularization techniques methods with preconditioning.
\end{abstract}

\keywords{Adam \and AdamW \and OASIS \and Regularization \and Weight Decay \and Optimization}

\section{Introduction}

In machine learning we consider unconstrained optimization problem
\begin{equation}
    \label{eq:general}
	\min_{w \in \mathbb{R}^d} f(w).
\end{equation}
Problems of the form \eqref{eq:general} cover a plethora of applications, including empirical risk minimization \citep{chapelle2000vicinal},
deep learning \citep{lecun2015deep}, and supervised learning \citep{cunningham2008supervised} tasks such as regularized least squares \citep{rifkin2007notes} or logistic regression \citep{shalev2014understanding}.

The classic base method for solving the optimization problem is gradient descent, but this minimization problem can be difficult to solve, particularly when the number of training samples, or problem dimension, is large.
In such cases, evaluating the full gradient on every iteration in the context of gradient descent becomes prohibitively expensive, especially considering that gradient descent often requires numerous iterations to converge.
In modern machine learning especially large problems represent the greatest interest.
For such cases stochastic gradient descent \citep{robbins1951stochastic} became popular solution.
Despite its simplicity, it proved itself to be an efficient and effective optimization method.
For a long time first-ordered methods were most popular approach of solving optimization problems.

Other way of solving the problem are methods with adaptive gradient \citep{hazan2007adaptive}. These methods posses several superiority over first-ordered methods.
Firstly, they have bigger potential of distributed solving, because first ordered methods spend majority of time on "communication".
Secondly, they are less sensitive to the choice of hyperparameters up to the point that hyperparameters can be set equal to one.
Lastly, this methods often simply show faster convergence on modern large optimization problems, especially this methods became applicable in neural networks solving \citep{kingma2014adam}.
Nowadays it is known that preconditioning methods often outperform other methods on modern large optimization problems \citep{goldberg2011oasis, kingma2014adam, zhang2018three, yao2021adahessian}.

Preconditioning methods refer to techniques that involve scaling the gradient of a problem by a specific matrix $D_t$, which enables the gradient to take into account the geometry of the problem. In the traditional case $D_t = (\nabla^2 f(w))^{-1}$, which corresponds Newton's method, however hessian is difficult to calculate and even more difficult to reverse, because of that some heuristics are used to replace the reversed hessian \citep{dennis1977quasi}. In OASIS \citep{jahani2021doubly} or AdaHessian \citep{yao2021adahessian} hessian is assumed to have diagonal dominance. In Adam \citep{kingma2014adam} gradient is simply normalized. This heuristics were proved to be effective. Generally, the step of preconditioned algorithms can be expressed as
\begin{equation}
    w_t = w_{t-1} - \eta \cdot D_t^{-1} g_t,
\end{equation}
where $\eta$ is a learning rate, $D_t$ is a preconditioning matrix, obtained through different heuristics, and $g_t$ is an unbiased stochastic gradient, i.e. $\mathbb{E}[g_t] = \nabla f(w)$.

We do not prescribe the exact method for obtaining $g_t$. In large-scale problems, stochastic gradient descent is a more efficient approach since computing the full gradient is prohibitively expensive.
While computing the Hessian is considerably more costly, so there exist various techniques for computing the preconditioning matrix $D_t$, and we delegate the specific choice of the preconditioning approach to the authors of individual methods.

Despite the various advantages offered by preconditioning methods, they are prone to overfitting. As a result, preconditioning methods in practice are almost always utilized with various methods to combat overfitting, with regularization being the primary such method employed.
Regularization is a powerful technique in machine learning that aims to prevent overfitting by adding additional constraints to the model.
It has been widely applied to various machine learning problems, including image classification \citep{zhu2017learning}, speech recognition \citep{zhou2017improved}, and natural language processing \citep{wu2022stgn}, and has shown its effectiveness in improving the generalization capability of neural networks \citep{girosi1995regularization}.

In methods with preconditioning appears to be several ways to include regularization.
We can include regularizer $r$ in $g_t$ calculation so it will be taken into consideration while calculating $D_t$. This method is equal to considering optimization problem 
\begin{equation}
	\min_{w \in \mathbb{R}^d} f(w) + r(w).
\end{equation}
Or we can include regularizer only on last step, decreasing norm of $w$ \citep{loshchilov2017decoupled}.
This way of regularization is called weight decay and surprisingly turns out to be more efficient in practical problems.
There are few other ways of considering regularizer which will be discussed further in the paper.

In general, our paper provides insight into comparison of different consideration ways of regularization is methods with preconditioning. Here, we provide a brief summary of our main contributions:

\begin{itemize}
    \item \textbf{Proof of preconditioned methods with weight decay convergence.}  We derive convergence guarantees for preconditioned methods considering assumptions of smoothness, strongly convex and PL-condition.

    \item \textbf{Research of the loss function.} Comparison of accuracy and loss function for AdamW and AdamL2. As a result we see that AdamW asymptotically converges to a non-zero value.

    \item \textbf{Competitive numerical results.} We investigate the empirical performance of Adam's variation including new one on a variety of standard machine learning tasks, including logistic regression.
\end{itemize}

\section{Regularizer consideration}

\if 0
\subsection{General preconditioning algorithm}

To begin with, we present the general formulation of preconditioned algorithms in \hyperref[alg:precond]{Algorithm 1}, and the general formulation of first-order algorithms in \hyperref[alg:first]{Algorithm 2}.
As mentioned briefly earlier, the key distinction of preconditioned methods is the application of the preconditioning matrix $D_t$ to scale the gradient, taking into account second-order information.

\begin{minipage}{0.47\textwidth}
    \begin{algorithm}[H]
        \caption{Method with preconditioning}
        \label{alg:precondgen}

        \begin{algorithmic}
                \Require{learning rate $\eta$, target function $f$}

                \While {$w$ not converged}
                \State $t = t+1$
                \State $g_t \gets$ stochastic gradient
                \State $D_t \gets$ preconditioning matrix
                \State $w_t \gets w_{t-1} - \eta \cdot D_t^{-1} g_t$
                \EndWhile
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.47\textwidth}
    \begin{algorithm}[H]
        \caption{First-ordered algorithm}
        \label{alg:first}
        \begin{algorithmic}
                \Require{learning rate $\eta$, target function $f$}

                \While {$w$ not converged}
                \State $t = t+1$
                \State $g_t \gets$ stochastic gradient
                \State $w_t \gets w_{t-1} - \eta \cdot g_t$
                \EndWhile
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

We do not prescribe the exact method for obtaining $g_t$. In large-scale problems, stochastic gradient descent is a more efficient approach since computing the full gradient is prohibitively expensive. While computing the Hessian is considerably more costly, so there exist various techniques for computing the preconditioning matrix $D_t$, and we delegate the specific choice of the preconditioning approach to the authors of individual methods.

\subsection{Three ways of regularizer consideration}

\fi

In preconditioned methods, there exist several techniques for incorporating regularization into the optimization process.
In this study, we consider three different approaches, which are illustrated in \hyperref[alg:precond]{Algorithm 1} using different colors.
In general, these methods can be characterized by the stage in which the regularization term is incorporated into the optimization process.

\begin{algorithm}[H]
    \caption{Different ways of regularization}
    \label{alg:precond}

    \begin{algorithmic}
            \Require{$\eta, f$}

            \While {$w$ not converged}
            \State $t = t+1$
            \State $g_t \gets$ stochastic gradient
            \State $\textcolor{blue}{g_t \gets g_t + \nabla r(w_t)}$ \hfill \textcolor{blue}{standart regularization}
            \State $D_t \gets$ preconditioning matrix
            \State $w_t \gets w_{t-1} - \eta \cdot D_t^{-1} \left(g_t +\textcolor{orange}{\nabla r(w_t)} \right) \textcolor{red}{ - \eta \cdot \nabla r(w_t)}$ \hfill \textcolor{orange}{hessian weight decay}, \textcolor{red}{weight decay}
            \EndWhile
    \end{algorithmic}
\end{algorithm}

To be more specific, the first regularization technique illustrated in \textcolor{blue}{blue} involves simply adding the regularization term to the objective function. This regularizer is included in the pseudo-gradient and factored into the calculation of $D_t$, and can be viewed as solving the following optimization problem:
\begin{equation*}
 \min_{w \in \mathbb{R}^d} F(w) = f(w) + r(w),
\end{equation*}
where $f(w)$ is the objective function and $r(w)$ is the regularization term. In essence, this approach involves applying the basic preconditioning method to the function $F$.

The second regularization technique, shown in \textcolor{orange}{orange}, is a novel approach.
Here, the regularization term is added before applying $D_t$, without affecting its computation.

The last regularization approach we consider is known as weight decay, illustrated by the color \textcolor{red}{red} in the \hyperref[alg:precond]{general scheme}.
This method only incorporates the regularizer during the algorithmic step, avoiding interference of regularization with the preconditioning stage.

Overall, it is important to carefully consider the impact of regularization when designing optimization algorithms, and we hope that our investigation of this techniques will prove useful to researchers in the field.

In our paper we mainly focus on weight decay regularization.
Though this technique is simpler than the others we consider, it can be quite effective in many cases.

\section{Weight decay regularization}

The most common regularization technique is $\ell_2$ regularization, defined as $r(w_t) = \frac{\lambda}{2} ||w_t||^2$, which yields $\nabla r (w_t) = \lambda \cdot w_t$.
In this case, the weight decay regularization achieves its name.
It only serves to reduce the norm of the weight vector, through the subtraction of $\eta \cdot w_t$ in the final step.

However in our theoretical evaluations we use regularizer in general form.
Thus, algorithmic step that we consider is 
\begin{equation}
w_{t+1} = w_t - \eta \cdot D_t^{-1} \nabla f(w_t) - \eta \cdot \nabla r(w_t),
\end{equation}
where $D_t$ is a preconditioning matrix, $\eta$ is a learning rate, $f$ is a target function and $r$ is a regularizer.

\subsection{Convergence speed of preconditioning methods}

We set ourselves a goal to estimate a convergence speed of methods with preconditioning with weight decay regularization.
Although step of methods with weight decay seems simple, it can be viewed in a rather unexpected way. We can put $D_t^{-1}$ out of brackets which gives
\begin{equation}
w_{t+1} = w_t - \eta \cdot D_t^{-1} (\nabla f(w_t) + D_t \nabla r(w_t)).
\end{equation}
That suggests the need to introduce a variable $\tilde{r}$ such that $\nabla \tilde{r}(w_t) = D_t \nabla r(w_t)$ and new target function $\tilde{F} = f + \tilde{r}$. New regularizer $\tilde{r}$ becomes adaptive, because both $r$ and $D_t$ depends on weight vector $w_t$.

The convergence speed is typically measured in terms of the number of iterations required to reach a certain level of error.
To obtain estimates on the number of iterations required to converge to a given error, we must impose certain assumptions on the function.

Throughout this work we assume that each f $:\mathbb{R}^d \rightarrow \mathbb{R}$ is convex, twice differentiable and L-smooth.
Additionally we imply a PL-condition to make another evaluation concerning speed of convergence.
This is formalized in the following assumptions.

\begin{assumption}{1}{Convexity}
	The function f is convex, i.e. $\forall y, x \in \mathbb{R}^d$
	\begin{equation*}
		f(y) \geq f(x) + \langle \nabla f(x), y-x \rangle
	\end{equation*}
        \label{ass:convex}
\end{assumption}

\begin{assumption}{2}{PL--condition}
	There exists $\mu > 0$, such that $||\nabla f(w) || \geq 2 \mu (f(w) - f^*)$, $\forall w \in \mathbb{R}^d$
        \label{ass:plcondition}
\end{assumption}

\begin{assumption}{3}{L-l-smoothness}
	The gradients of F are L-Lipschitz continuous $\forall w \in \mathbb{R}^d$, i.e. there exists a constant $L > 0$ such that $\forall x, y \in \mathbb{R}^d$,
	\begin{equation*}
		f(x) \leq f(y) + \langle \nabla f(y), x-y \rangle + \frac{L}{2} ||x - y||^2
	\end{equation*}
    \label{ass:Lsmoothness}

    The gradient of r is l-Lipschitz continuous $\forall w \in \mathbb{R}^d$, i.e. there exists a constant $l > 0$ such that $\forall x, y \in \mathbb{R}^d$,
	\begin{equation*}
		r(x) \leq r(y) + \langle \nabla r(y), x-y \rangle + \frac{l}{2} ||x - y||^2
	\end{equation*}
        \label{ass:lsmoothness}
\end{assumption}

Also we must introduce restrictions on preconditioner $D_t$
\begin{equation}
\alpha I \preccurlyeq D_t \preccurlyeq \Gamma I \Leftrightarrow \frac{I}{\alpha} \preccurlyeq D_t^{-1} \preccurlyeq \frac{I}{\Gamma}
\end{equation}

Using introduced assumptions we prooved convergence of methods with preconditioning with weight decay regularization in general form. Our results are framed in \hyperref[theor:1]{Theorem 1} and \hyperref[theor:2]{Theorem 2}

\begin{theorem}
Suppose the Assumption \hyperref[ass:convex]{1}, \hyperref[ass:Lsmoothness]{2} hold, let $\varepsilon > 0$ and let the step-size satisfy
\begin{equation*}
    \eta < \frac{2 \alpha}{L + l \cdot \alpha}
\end{equation*}
Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem \eqref{eq:general} can be bounded by
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}
\label{theor:1}
\end{theorem}

\begin{theorem}
    Suppose the Assumption \hyperref[ass:convex]{1}, \hyperref[ass:Lsmoothness]{2}, \hyperref[ass:plcondition]{3} hold, let $\varepsilon > 0$ and let the step-size satisfy
    \begin{equation*}
        \eta \leq \frac{2 \alpha}{\tilde{L}}
    \end{equation*}
    Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem \eqref{eq:general} can be bounded by
    \begin{equation*}
        T =  \mathcal{O}\left( \frac{\ln \frac{\Delta_0}{\epsilon}}{2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha})} \right)
    \end{equation*}
    \label{theor:2}
\end{theorem}

\subsection{Convergence trajectory of preconditioning methods}

 In the previous subsection we have proved convergence of preconditioned methods, however we have pointed out above that methods with weight decay does not converge to the initial optimized function $F = f + r$, but rather to a new function $\tilde{F} = f + \tilde{r}$.

We will consider two algorithms OASIS and Adam, and its variations. Their main difference is in the calculation of the pseudo hessian.
In Adam, the Hessian is a diagonal matrix consisting of squares of derivatives, in OASIS we have a stochastic Hessian, which is calculated through a random variable from the Randemacher distribution.

We framed three methods of regularization for Adam and OASIS in \hyperref[alg:genAdam]{Algorithm 2} and \hyperref[alg:OASIS]{Algorithm 3} respectively.
Results of our computational experiments are framed in \hyperref[fig:adams_error]{Figure 1} and \hyperref[fig:oasis_error]{Figure 2}.
In can be seen that methods with weight decay converges only by special criterion $\nabla f + D_t \nabla r$.

\begin{algorithm}[h]

            \caption{Adam}\label{alg:genAdam}

            \begin{algorithmic}
            \small{
            \Require{$\eta, \beta_1, \beta_2, \epsilon, f, r$}
            %\State $m_0 = 0$ -- 1-st moment vector
            %\State$v_0 = 0$ -- 2-nd moment vector
            \While {$\theta$ not converged}
            \State $t = t+1$
            \State $g_t = \nabla f(w_{t-1}) + $ \textcolor{blue}{$\nabla r(w_{t-1})$}\hfill \textcolor{blue}{AdamL2}
            \State $m_t = \beta_1 \cdot m_{t-1} + (1 - \beta_1) \cdot g_t$
            \State $v_t = \beta_2 \cdot v_{t-1} + (1 - \beta_2) \cdot g_t^2$
            \State $\hat{m_t} = \frac{m_t}{1-\beta_1^t} +$ \textcolor{orange}{$\nabla r(w_{t-1})$} \hfill \textcolor{orange}{AdamWH}
            \State $\hat{v_t} = \frac{v_t}{1-\beta_2^t}$ 
            \State $w_t = w_{t-1} - \eta \cdot \frac{\hat{m_t}}{\sqrt{v_t} + \epsilon} - $ \textcolor{red}{$\eta \nabla r(w_{t-1})$ } \hfill \textcolor{red}{AdamW}
            \EndWhile
            }
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
\caption{OASIS}\label{alg:OASIS}
\begin{algorithmic}
    \Require{$w_0, \eta_0, D_0, \theta_0 = + \infty$}    
    \State $w_1 = w_0 - \eta \hat{D_0}^{-1} \nabla f(w_0)$

    \For{$k = 1, 2, ...$}
    \State $g_k = \nabla f(w_k) +$ \textcolor{blue}{$\nabla r(w_{t-1})$}\hfill \textcolor{blue}{OASISL2} 
    \State $D_k = \beta D_{k-1} + (1-\beta_2) \cdot diag\left( z_k \odot \nabla^2 \left(f(w_k) + \textcolor{orange}{r(w_k)} \right) z_k \right)$ \hfill \textcolor{orange}{OASISWH}
    \State $(\hat{D_k})_{ii} = max \{|D_k|_{i, i} ; \alpha \}$, $\forall i = \overline{1, d}$
    \State $\eta_k = min \{ \sqrt{1 + \theta_{k-1}} \cdot \eta_{k-1}; \frac{||w_k - w_{k-1}||_{\hat{D_k}}}{2 ||\nabla f(w_k) - \nabla f(w_{k-1}) ||_{\hat{D_k}}^* } \}$
    \State $w_{k+1} = w_k - \eta_k g_k D_k^{-1}- $ \textcolor{red}{$\eta \nabla r(w_{t-1})$ } \hfill \textcolor{red}{OASISW} 
    \State $\theta_k = \frac{\eta_k}{\eta_{k-1}}$
    \EndFor

\end{algorithmic}
\end{algorithm}

Thus, method converges to another solution,

\begin{minipage}{0.5\textwidth}
\begin{figure}[H]
\centering
    \includegraphics[width=\textwidth]{pictures/mushrooms/adams_error.png}
    \caption{Adam on dataset mushrooms}
        \label{fig:adams_error}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}{0.5\textwidth}
\begin{figure}[H]
\centering
    \includegraphics[width=\textwidth]{pictures/mushrooms/oasis_error.png}
    \caption{OASIS on dataset mushrooms}
        \label{fig:oasis_error}
\end{figure}
\label{fig:error}
\end{minipage}

\begin{figure}[H]
\centering
    \includegraphics[width=0.9\textwidth]{pictures/mushrooms/main_adam.png}
    \caption{Dataset: mushrooms}

        \label{fig:main_mushrooms_adam}
\end{figure}

\begin{figure}[H]
\centering
    \includegraphics[width=0.9\textwidth]{pictures/mushrooms/main_mushrooms.png}
    \caption{Dataset: mushrooms}

        \label{fig:main_mushrooms}
\end{figure}

\begin{figure}[H]
\centering
    \includegraphics[width=0.9\textwidth]{pictures/wine/main_wine.png}
    \caption{Dataset: wine}

        \label{fig:main_wine}
\end{figure}

\subsection{Computational experiment}

\bibliographystyle{unsrtnat}
\bibliography{refs}

\newpage
\section{Appendix}
\begin{theorem}
Suppose the Assumption \hyperref[ass:convex]{1}, \hyperref[ass:Lsmoothness]{2} and let $\varepsilon > 0$ and let the step-size satisfy
\begin{equation*}
    \eta < \frac{2 \alpha}{L + l \cdot \alpha} 
\end{equation*}
Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem \eqref{eq:general} can be bounded by
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}

@@

Let's write \ref{ass:convex} for step t and $t+1$:

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \nabla f(w_t), w_{t+1} - w_t \rangle + \frac{L}{2}||w_{t+1} - w_t ||^2,
\end{equation*}
Okay, by definition for our algorithm we have:

\begin{equation*}
w_{t+1} - w_t = -\eta D_t^{-1} \nabla f(w_t) - \eta \nabla r(w_t),
\end{equation*}

from the previous expression , we express the gradient of the function

\begin{equation*}
\nabla f(w_t) = \frac{1}{\eta} D^t(w_t - w_{t+1}) - D^t \nabla r(w_t),
\end{equation*}

replace $\nabla f(w_t)$ and by definition of matrix $D_t$, $I \leq \frac{D_t}{\alpha}$
\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \frac{1}{\eta}D_t(w_t - w_{t+1}) - D_t\nabla r(w_t), w_{t+1} - w_t \rangle + \frac{L}{2 \alpha} ||w_{t+1} - w_t||_{D_t}^2,
\end{equation*}

now let's bring it together
\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left(\frac{L}{2 \alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 - \langle D_t \nabla r(w_t), w_{t+1} - w_t \rangle,
\end{equation*}

define new regularization function $\tilde{r} : \nabla \tilde{r} = D_t \nabla r(w_t)$.

then rewrite step using the variable and assumption \ref{ass:lsmoothness}
\begin{equation*}
    \tilde{r}(w_{t+1}) \leq \tilde{r}(w_t) + \langle \tilde{r}(w_t), w_{t+1} - w_t \rangle + \frac{l}{2} (w_{t+1} - w_t)^T D_t (w_{t+1} - w_t),
\end{equation*}

let's replace the old regularization function with a new one

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left( \frac{L}{2\alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 + \tilde{r}(w_t) - \tilde{r}(w_{t+1}) + \frac{l}{2}||w_{t+1}-w_t||_{D_t}^2,
\end{equation*}
